Negative Perceptions of Outsourcing to Artificial Intelligence

### 2025-04-30 appendix.qmd
Edit manuscript

We recruited a convenience sample of 500 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of 
`r nrow(study1_sample)` participants (`r sum(study1_sample$gender == "Female")`
female; `r sum(study1_sample$gender == "Male")` male;
`r sum(study1_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study1_sample$gender, "Prefer") | is.na(study1_sample$gender))`
undisclosed gender; mean age = `r round(mean(study1_sample$age), 2)` years).
`r round(mean(study1_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-pilotstudy2).

@@

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was (see @suppfig-chatgpt-pilotstudy2).

@@

::: {#suppfig-chatgpt-study1 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study2)
```

Responses to the questions about ChatGPT in Study 1.

@@

::: {#suppfig-chatgpt-study2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study3)
```

Responses to the questions about ChatGPT in Study 2.

@@

::: {#suppfig-chatgpt-study3 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study4)
```

Responses to the questions about ChatGPT in Study 3.

@@

::: {#suppfig-chatgpt-study4 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study5)
```

Responses to the questions about ChatGPT in Study 4.

@@

::: {#suppfig-chatgpt-study5 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study6)
```

Responses to the questions about ChatGPT in Study 5.

@@

::: {#suppfig-chatgpt-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study1)
```

Responses to the questions about ChatGPT in the second pilot study.

>>>>

We recruited a convenience sample of 500 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of 
`r nrow(study1_sample)` participants (`r sum(study1_sample$gender == "Female")`
female; `r sum(study1_sample$gender == "Male")` male;
`r sum(study1_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study1_sample$gender, "Prefer") | is.na(study1_sample$gender))`
undisclosed gender; mean age = `r round(mean(study1_sample$age), 2)` years).
`r round(mean(study1_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

@@

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was.

@@

@@

@@

@@

@@

@@

### 2025-04-30 manuscript.qmd
Edit manuscript

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: All data and original code can be found on GitHub at <https://github.com/ScottClaessens/outsourcing>.
    related-report: ~
    conflict-of-interest: The authors have no conflicts of interest to disclose.
    financial-support: This work was generously supported by funding from the Horizon Europe UK Guarantee via the UKRI (EP/Y00440X/1) and a Philip Leverhulme Prize (PLP-2021-095) awarded to JACE.
    gratitude: ~
    authorship-agreements: ~
abstract: >
  As artificial intelligence (AI) tools become increasingly integrated into 
  daily life, people are beginning to outsource not only professional tasks but 
  also socio-relational ones. Large language models like ChatGPT can generate 
  wedding vows, speeches, and personal messages, raising questions about how 
  individuals who use AI for such tasks are perceived by others. In this paper,
  we conduct five pre-registered studies with British participants (N = 3,649)
  to understand how people view those who outsource tasks to AI, and how this
  depends on how socio-relational the task is, whether AI is used as a tool or
  fully delegated to, and the acknowledgment of the AI use. We find negative
  perceptions of outsourcing, particularly for socio-relational tasks. We show
  that outsourcing makes us think more negatively about not only the person and
  their motivations, but also the outsourced work itself. Moreover, we provide
  insight into why this occurs: the reduced effort from outsourcing
  socio-relational tasks to AI signals that the output is less authentically 
  one's own and that the person cares less about the task. Our research 
  highlights the way that AI use shapes our perceptions of people, raising key
  philosophical questions about efficiency, authenticity, and social ties in a
  world filled with AI-mediated interactions.
keywords: [artificial intelligence, person perception, outsourcing, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-pdf: default
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

@@

Second, we need to know whether the *type* of task that people are outsourcing 
matters. One might expect outsourcing to be perceived negatively regardless of 
the type of task being outsourced -- if effort is generally moralised, then the 
domain in which it is expended (or not) should have little impact. However, 
there are reasons to expect differences between social tasks like writing vows 
and non-social tasks like writing computer code. We know that different norms,
standards, and expectations can be applied to social and non-social tasks and 
exchanges [e.g. @Fiske1992; @Heider1958; @Malle2022]. Moreover, from a 
philosophical perspective, it often matters not only *whether* something is 
done, but *how* it is done [@Stohr2006]. An apology is not just about hearing 
someone say "I am sorry", but seeing genuine regret; a love letter is not just 
about hearing someone say "I love you", but seeing depth of emotion; and a 
bereavement letter is not just about hearing someone say "I am sorry for your 
loss", but seeing an understanding for the powerful human experience of loss.
There is, perhaps especially for social tasks, value not only in the outcome of 
doing something, but the *process* too [@Goodman2010]. To understand any 
potential negative effects of outsourcing to AI, we must therefore look at a 
broad range of non-social and social tasks, rather than draw broad conclusions
based on a few use cases.

@@

We recruited a convenience sample of 800 participants from the United Kingdom 
through the online platform Prolific (<https://www.prolific.com/>). After 
excluding participants who failed our pre-treatment attention check, we were 
left with a final sample of `r nrow(study2_sample)` participants 
(`r sum(study2_sample$gender == "Female")` female; 
`r sum(study2_sample$gender == "Male")` male;
`r sum(study2_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study2_sample$gender, "Prefer") | is.na(study2_sample$gender))`
undisclosed gender; mean age = `r round(mean(study2_sample$age), 2)` years).
`r round(mean(study2_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-study1).

@@

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was (see @suppfig-chatgpt-study1).

@@

We recruited a convenience sample of 800 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study3_sample)` 
participants (`r sum(study3_sample$gender == "Female")` female; 
`r sum(study3_sample$gender == "Male")` male;
`r sum(study3_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study3_sample$gender, "Prefer") | is.na(study3_sample$gender))`
undisclosed gender; mean age = `r round(mean(study3_sample$age), 2)` years).
`r round(mean(study3_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-study2).

@@

We used the same power estimate from Study 1 to determine our target sample size
of *n* = 750 (150 participants in each of five conditions). We recruited a 
convenience sample of 802 participants from the United Kingdom through Prolific.
After excluding participants who failed our pre-treatment attention check, we
were left with a final sample of `r nrow(study4_sample)` 
participants (`r sum(study4_sample$gender == "Female")` female; 
`r sum(study4_sample$gender == "Male")` male;
`r sum(study4_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study4_sample$gender, "Prefer") | is.na(study4_sample$gender))`
undisclosed gender; mean age = `r round(mean(study4_sample$age), 2)` years).
`r round(mean(study4_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-study3).

@@

We used the same power estimate from Study 1 to determine our target sample size
of *n* = 750 (150 participants in each of five conditions). We recruited a 
convenience sample of 800 participants from the United Kingdom through Prolific. 
After excluding participants who failed our pre-treatment attention check, we 
were left with a final sample of `r nrow(study5_sample)` 
participants (`r sum(study5_sample$gender == "Female")` female; 
`r sum(study5_sample$gender == "Male")` male;
`r sum(study5_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study5_sample$gender, "Prefer") | is.na(study5_sample$gender))`
undisclosed gender; mean age = `r round(mean(study5_sample$age), 2)` years).
`r round(mean(study5_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-study4).

@@

We recruited a convenience sample of 651 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study6_sample)` 
participants (`r sum(study6_sample$gender == "Female")` female; 
`r sum(study6_sample$gender == "Male")` male;
`r sum(study6_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study6_sample$gender, "Prefer") | is.na(study6_sample$gender))`
undisclosed gender; mean age = `r round(mean(study6_sample$age), 2)` years).
`r round(mean(study6_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used AI tools like ChatGPT before (see 
@suppfig-chatgpt-study5).

@@

Our findings cohere with the philosophical idea that there is value in *how* a 
task was done, and not merely *whether* it was done [@Goodman2010; @Stohr2006]. 
For many socio-relational tasks, it might seem that part of the constitutive 
action is the *process* by which it occurs: an apology that does not contain a 
genuine reflection and commitment to do better, rather than just the words "I am 
sorry", might not seem to be an apology at all. In contrast, for many of the 
non-social tasks, it is easier to distinguish the importance of the process from 
the outcome. In this way, our work suggests that people rarely adopt a purely 
utilitarian perspective in which outcomes are the sole determinant 
[@Everett2020; @Kahane2018]. Instead, their judgments cohere more with ideas 
from virtue ethics about the importance of *doing* [@Stohr2006]. Outsourcing to 
AI -- especially for social tasks –- may allow us to produce similar outputs, 
but by severing the outcome from the practice of doing, it may risk the 
development and maintenance of our human virtues [@Vallor2015].

@@

>>>>

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed, and may differ from the final version.
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: >
  As artificial intelligence (AI) tools become increasingly integrated into 
  daily life, people are beginning to outsource not only professional tasks but 
  also socio-relational ones. Large language models like ChatGPT can generate 
  wedding vows, speeches, and personal messages, raising questions about how 
  individuals who use AI for such tasks are perceived by others. In this paper,
  we conduct five pre-registered studies with British participants (N = 3,649)
  to understand how people view those who outsource tasks to AI, and how this
  depends on how socio-relational the task is, whether AI is used as a tool or
  fully delegated to, and the acknowledgment of the AI use. We find negative
  perceptions of outsourcing, particularly for socio-relational tasks. We show
  that outsourcing makes us think more negatively about not only the person and
  their motivations, but also the outsourced work itself. Moreover, we provide
  insight into why this occurs: the reduced effort from outsourcing
  socio-relational tasks to AI signals that the output is less authentically 
  one's own and that the person cares less about the task. Our research 
  highlights the way that AI use shapes our perceptions of people, raising key
  philosophical questions about efficiency, authenticity, and social ties in a
  world filled with AI-mediated interactions.
keywords: [artificial intelligence, person perception, outsourcing, effort, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-pdf: default
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

@@

Second, we need to know whether the *type* of task that people are outsourcing 
matters. One might expect outsourcing to be perceived negatively regardless of 
the type of task being outsourced -- if effort is generally moralised, then the 
domain in which it is expended (or not) should have little impact. However, 
there are reasons to expect differences between social tasks like writing vows 
and non-social tasks like writing computer code. We know that different norms,
standards, and expectations can be applied to social and non-social tasks and 
exchanges [e.g. @Fiske1992; @Heider1958; @Malle2022]. Moreover, from a 
philosophical perspective, it often matters not only *whether* something is 
done, but *how* it is done [@Aristotle2009; @Hursthouse2023; @Stohr2006]. An 
apology is not just about hearing someone say "I am sorry", but seeing genuine 
regret; a love letter is not just about hearing someone say "I love you", but 
seeing depth of emotion; and a bereavement letter is not just about hearing 
someone say "I am sorry for your loss", but seeing an understanding for the 
powerful human experience of loss. There is, perhaps especially for social 
tasks, value not only in the outcome of doing something, but the *process* too 
[@Goodman2010]. To understand any potential negative effects of outsourcing to 
AI, we must therefore look at a broad range of non-social and social tasks, 
rather than draw broad conclusions based on a few use cases.

@@

We recruited a convenience sample of 800 participants from the United Kingdom 
through the online platform Prolific (<https://www.prolific.com/>). After 
excluding participants who failed our pre-treatment attention check, we were 
left with a final sample of `r nrow(study2_sample)` participants 
(`r sum(study2_sample$gender == "Female")` female; 
`r sum(study2_sample$gender == "Male")` male;
`r sum(study2_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study2_sample$gender, "Prefer") | is.na(study2_sample$gender))`
undisclosed gender; mean age = `r round(mean(study2_sample$age), 2)` years).
`r round(mean(study2_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

@@

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was.

@@

We recruited a convenience sample of 800 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study3_sample)` 
participants (`r sum(study3_sample$gender == "Female")` female; 
`r sum(study3_sample$gender == "Male")` male;
`r sum(study3_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study3_sample$gender, "Prefer") | is.na(study3_sample$gender))`
undisclosed gender; mean age = `r round(mean(study3_sample$age), 2)` years).
`r round(mean(study3_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

@@

We used the same power estimate from Study 1 to determine our target sample size
of *n* = 750 (150 participants in each of five conditions). We recruited a 
convenience sample of 802 participants from the United Kingdom through Prolific.
After excluding participants who failed our pre-treatment attention check, we
were left with a final sample of `r nrow(study4_sample)` 
participants (`r sum(study4_sample$gender == "Female")` female; 
`r sum(study4_sample$gender == "Male")` male;
`r sum(study4_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study4_sample$gender, "Prefer") | is.na(study4_sample$gender))`
undisclosed gender; mean age = `r round(mean(study4_sample$age), 2)` years).
`r round(mean(study4_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

@@

We used the same power estimate from Study 1 to determine our target sample size
of *n* = 750 (150 participants in each of five conditions). We recruited a 
convenience sample of 800 participants from the United Kingdom through Prolific. 
After excluding participants who failed our pre-treatment attention check, we 
were left with a final sample of `r nrow(study5_sample)` 
participants (`r sum(study5_sample$gender == "Female")` female; 
`r sum(study5_sample$gender == "Male")` male;
`r sum(study5_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study5_sample$gender, "Prefer") | is.na(study5_sample$gender))`
undisclosed gender; mean age = `r round(mean(study5_sample$age), 2)` years).
`r round(mean(study5_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

@@

We recruited a convenience sample of 651 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study6_sample)` 
participants (`r sum(study6_sample$gender == "Female")` female; 
`r sum(study6_sample$gender == "Male")` male;
`r sum(study6_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study6_sample$gender, "Prefer") | is.na(study6_sample$gender))`
undisclosed gender; mean age = `r round(mean(study6_sample$age), 2)` years).
`r round(mean(study6_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used AI tools like ChatGPT before.

@@

Our findings cohere with the philosophical idea that there is value in *how* a 
task was done, and not merely *whether* it was done [@Aristotle2009; 
@Goodman2010; @Hursthouse2023; @Stohr2006]. For many socio-relational tasks, it 
might seem that part of the constitutive action is the *process* by which it 
occurs: an apology that does not contain a genuine reflection and commitment to 
do better, rather than just the words "I am sorry", might not seem to be an 
apology at all. In contrast, for many of the non-social tasks, it is easier to 
distinguish the importance of the process from the outcome. In this way, our 
work suggests that people rarely adopt a purely utilitarian perspective in which 
outcomes are the sole determinant [@Everett2020; @Kahane2018]. Instead, their 
judgments cohere more with ideas from virtue ethics about the importance of 
*doing* [@Hursthouse2023; @Stohr2006]. Outsourcing to AI -- especially for 
social tasks –- may allow us to produce similar outputs, but by severing the 
outcome from the practice of doing, it may risk the development and maintenance
of our human virtues [@Vallor2015; @Vallor2024].

@@

# Acknowledgements

This work was generously supported by funding from the Horizon Europe UK 
Guarantee via the UKRI (EP/Y00440X/1) and a Philip Leverhulme Prize 
(PLP-2021-095) awarded to JACE.

# Data and Code Availability

All data and original code can be found on GitHub at
<https://github.com/ScottClaessens/outsourcing>.

# Statement of Interests

The authors have no conflicts of interest to disclose.
